### 2025-01-24 report/msds_data_mining_project_report.tex
add trimmed versions of sigconf acmart latex template and bibtex that are working when compiling for pdf

\documentclass[sigconf, authorversion, nonacm]{acmart}

\settopmatter{printacmref=false}

\citestyle{acmauthoryear}

\begin{document}

\title{Higher Education Synthesized}
\subtitle{Extending Higher Education Data with Synthetic Data at the Onset of the Demographic Cliff}

\renewcommand{\shortauthors}{Leakakos}

\begin{abstract}
One to two paragraph executive summary of project and presentation.

One to two paragraph executive summary of project and presentation.
\end{abstract}

\maketitle

\section{Introduction}
What is the problem?
Why is it important?
Limitations of existing solutions
What is your (potential) contribution?

\section{Related Work}
What has been done already?
Group by category
    - E.g., tools, methods, results
How does your project build upon prior work?
    - E.g., utilize, compare, improve, new knowledge

\cite{DBLP:journals/corr/abs-1907-00503}

\cite{decristofaro2024syntheticdatamethodsuse}

\section{Proposed Work}
Dataset, tools
Main tasks
    - E.g., statistical analysis, visualization
    - E.g., data preprocessing, warehousing
    - E.g., specific questions, patterns to explore or model

\section{Evaluation}
Evaluation metrics
    - Effectiveness
    - Efficiency
Experimental setup
Methods to compare

\section{Discussion}
Project timeline
    - When to finish
    - Current status
    - Determine if you need to do less or can do more in the given time
Potential challenges
Alternative approaches and backup plan

\section{Conclusion}
Conclusion
Project summary
Key findings (when finished)
Future work (when finished)

\bibliographystyle{ACM-Reference-Format}
\bibliography{msds_data_mining_project_report}

\end{document}
\endinput