### 2025-02-07 report/msds_data_mining_project_report.tex
fill out first pass on related work section and associated bibliography entries

\documentclass[sigconf, authorversion, nonacm]{acmart}

\settopmatter{printacmref=false} % remove acm info that clutters things up for this project
\settopmatter{printfolios=true} % add page numbers

\citestyle{acmauthoryear}

\begin{document}

\title{Higher Education Synthesized}
\subtitle{Extending Higher Education Data with Synthetic Data at the Onset of the Demographic Cliff}

\renewcommand{\shortauthors}{Leakakos}

\begin{abstract}
    American higher education is coming to an impasse. Colleges and universities have seen enrollments grow as larger generations of Americans made their way through postsecondary education. Schools increased their academic offerings and student life activities. Faculty and staff members filled the needs for supporting increasing student enrollments. But there was an issue hiding behind all of this growth: what happens when student enrollments systematically fall?

    The United States has seen declines in high school graduates and, therefore, reduced college applicants. Depending on who you listen to, this ranges from a demographic cliff ushering in an emergency in higher education to a slower decline in higher education enrollments that will play out over years. It is unclear what effect this will have on the near future of higher education. One expected consequence of these declines is that we will see smaller amounts of higher education data relative to previous years.

    This project focuses on synthesizing Integrated Postsecondary Education Data System data from the National Center for Education Statistics. This data covers a vast range of information about higher education, and, while we will not look at student-level data --- one of the key areas for synthetic data after the demographic cliff hits --- we will start by generating synthetic data based on subsets of institution-related data. The goal is to show how we might address the diminishing data after the demographic cliff in order to continue higher education data analysis that relies on larger datasets.
\end{abstract}

\maketitle

\section{Introduction}

    American higher education saw growing enrollments throughout the latter half of the 1900s and moving into the 200s. This resulted in increases in the number of higher education institutions (HEIs) in the country, including public and private colleges and universities. These HEIs span from major research institutions down to small liberal arts schools. This has also resulted in increases in the number of faculty and staff to support growing enrollments. The general trend in higher education was growth.

    The issue is that this growth in enrollments was not to last. For years now, we have seen predicted and then realized declines in high school graduates along with the respective drop in higher education enrollments. Higher education's future was based on growth, but then the key aspect of that growth --- more students than in previous years --- fell away. Estimates for what comes next vary. Some see a demographic cliff and a looming emergency in which declining enrollments result in catastrophic cuts at and closures of HEIs. It is true that we have seen increasing numbers of HEI closures in recent years ~\cite{hechingercollegeclosures}. But others see the trend as a more gradual decline, giving higher education time to adapt and shrink accordingly. \cite{insidehighereddemographiccliff}

    There are many interesting research questions related to the future of higher education, appropriate and responsible responses to enrollment declines, and other aspects of this impending reality. There are also questions related to the shifting of costs from public funding to individual students, the shift in emphasis to higher education as workforce preparation, the modern publishing culture in higher education and how that affects the student experience, and more.

    But, this project will instead focus on a different aspect of declining higher education enrollments: reduced data. As student enrollments drop and HEIs close or shrink, we are likely to see reduced amounts of higher education data available for analysis, prediction, evaluation, and more. Not all analyses in higher education need larger amounts of data, but we want to explore how we could supplement real data with synthetic data in cases where we need more overall data. The risk is that, without sufficient data, we may not be able to support students and HEIs as well as we could with larger amounts of data.

    We will take a look at Integrated Postsecondary Education Data System (IPEDS) data from the National Center for Education Statistics (NCES), housed in the U.S. Department of Education's Institute of Education Sciences (IES). NCES gathers, prepares, and makes available data along with providing analysis on education data. IPEDS is a subset of NCES data that focuses on postsecondary or higher education data.

    We will focus on a subset of IPEDS data related to institutions, enrollments, and graduation rates. For each of these three domains, IPEDS offers multiple files covering different information, and it provides these for multiple years. Our goal will be to identify one or more analysis or modeling datasets that we can pull out of a larger extract-transform-load (ETL) pipeline with multiple repeated steps. We will focus on pulling IPEDS data, loading it into a relational database in its raw form, preprocessing it in the database with a series of automated processes, storing the analysis or modeling datasets in the database, and then using those datasets to synthesize new data.

   To address the issue of the amount of data available to us, we will leverage generative modeling to produce synthetic tabular data. The goal is to build a system that can demonstrate how we might leverage synthetic higher education data to fill in datasets in the future as they potentially shrink due to the issues described above.

    We are focusing on institution-level data from IPEDS, but future iterations of this work could pivot to generating synthetic student-level data.

    Synthetic data is ``artificially generated [data] that resemble[s] the actual data -- more precisely, having similar statistical properties'' ~\cite[p. 1]{decristofaro2024syntheticdatamethodsuse}. If we generate synthetic IPEDS data that shares the same properties as the real data, then we can combine the real and synthetic data, or, alternatively, we can create an indefinite amount of synthetic data, to increase the amount of data we have to work with.

    Building and training models to generate synthetic tabular data -- tabular meaning data in table form with rows and columns -- is notoriously difficult. We will leverage existing models and research for the synthetic generation aspect of the pipeline in particular. We will lean on the research in the Modeling Tabular Data Using Conditional GAN paper ~\cite{DBLP:journals/corr/abs-1907-00503} and the associated Python libraries they make available at the Synthetic Data Vault Project ~\cite{sdv}.

    While we will not dive too deep into the working of the generative models, it can be useful to understand the complexity in these and why we are leveraging existing implementations. The research paper focuses on their conditional tabular generative adversarial network (CTGAN) approach to generative modeling. This paper provides some examples of the difficulty of modeling for tabular data. Section 3 speaks to some of these issues:

    \begin{itemize}
        \item Mixed data types: tabular data can have features with continuous, nominal, ordinal, binary, or other data. Generative modeling for these datasets require handling mixtures of data types in one model.
        \item Non-Gaussian distributions: continuous data can be non-Gaussian, so attempts to standardize or min-max continuous features can result in issues such as vanishing or exploding gradients.
        \item Multimodal distributions: continuous features may have more than one mode, resulting in the need to model more than one mode with approaches such as kernel density estimation and Gaussian mixture models.
        \item Imbalance in categories for categorical features: categories that have smaller representation may not be modeled as accurately as those with larger representation, resulting in the need for extra steps to handle this disparity. \cite{DBLP:journals/corr/abs-1907-00503}
    \end{itemize}

    We do not need to go into the inner workings of these issues. Instead, they are meant to provide a snippet of the difficulties we would need to handle if building our own generative models. In order to remain focused on the pipeline in this project, we will, as said above, leverage existing generative modeling libraries available through the Synthetic Data Vault Project ~\cite{sdv}.

    Finally, the ultimate goal is to glean insights from the IPEDS data, real or synthesized. The synthetic data will ideally give us the same overall insights as the real data. For this task, we can turn to traditional data mining techniques: classification, clustering, association mining, rule mining, or others. If we can determine trends now, we may be able to utilize synthetic data in the future with dwindling real data to confirm continued trends or to identify changes in those trends.

    To summarize, this project will work with IPEDS data with its focus on higher education and will explore how we can synthesize more IPEDS data to address potential future lack-of-data issues due to declining higher education enrollments, and we will use these data to tackle standard data mining questions related but focused on higher education data.

\section{Related Work}
    My literature review did not return as much related to data mining in higher education as I expected to see, and I was unable to find any research closely related to what we explore in this project. The bulk of what I found is related to different evaluation methods for student learning. Examples are looking at adaptive learning --- learning that dynamically adjusts to students based on different input metrics --- and performance prediction --- trying to determine how a student will end up performing by, for example, graduation. While these are interesting, they do not focus on the type of data that we see in IPEDS.

    Note that some of these papers were already referenced in the introduction above.

    To start with, ~\cite{sdp} by the Harvard Strategic Data Project (SDP) fellowship program speaking about the importance of analytics in higher education data. There are larger questions that this paper raises about why higher education appears behind other areas in terms of leveraging data and more advanced analytics.

    Another example of an interesting paper that I want to highlight is ~\cite{Hassna02102023} that explores utilizing big data and analytics for higher education data. The catch is that this paper is from 2022, and it is talking about introducing big data and analytics in higher education, mirroring some of what we see from the SDP paper above in hinting that higher education may be behind other areas in terms of utilizing data. But, during the literature review, I checked out big data and data mining and related terms in other domains, and I did not tend to see papers talking about how big data and analytics have yet to be realized in those domains.

    Next is a handful of papers characterizing what I mostly found: research into specific uses of analytics and mining to hone in on an aspect of student evaluation. Since our focus on this project is on institution-level information with IPEDS data, we are zooming out quite a bit from what these papers cover, but these are useful in terms of getting bearings for research on higher education data mining, though none of these papers are overly useful for the specific use case in this project.

    First is ~\cite{studentbehavioranalytics} that talks about using higher education big data to evaluate student behavior. This paper repeats a similar claim to ~\cite{Hassna02102023} above in that higher education is missing work in big data.

    \cite{Okewu10112021} explores using artificial neural networks for higher education data mining. It provides a review of studies from 2010 through 2018 and related to education data mining focused on adaptive learning. Adaptive learning is the use of learner data to customize learning dynamically to those learners. So, we are looking at yet another narrower use case for a specific type of higher education learner data. This paper again mirrors some of the sentiments from earlier papers in that it hints that higher education data usage, and big data usage in particular, should be further along than where it is currently at (as of the paper's publishing in 2021), closer to where we see big data usage in other domains, but it is not.

    \cite{10.3389/fpsyg.2021.698490} investigates education data mining techniques for student performance prediction. There is not a ton more to add here aside from it being another narrower use case looking at student-level data, not the larger institution-level data that we explore in this project, but it is yet another example to learn from in terms of research using higher education data.

    \cite{ALDOWAH201913} digs into education data mining, similar to the last paper, but this time focused on learning analytics. It reviews studies in different domains related to learning analytics. The key takeaway is that it is yet another study that looks into different outcomes within specific areas of higher education data, but it is not close enough to what we are exploring in this project to lean on.

    Finally, ~\cite{https://doi.org/10.1155/2022/8924028} looks at predicting higher education student performance, typically at graduation, and it evaluates different predictive methods. Again, this is interesting for reviewing some current literature, but it will not apply to the work we do in this project.

    We will also include a quick review of a few papers related to synthetic data and synthetic tabular data generative modeling. We will utilize synthetic data, but we are not building it from scratch, so we will not dwell on the deep learning implementation details of these papers.

    \cite{decristofaro2024syntheticdatamethodsuse} provides an overview of synthetic data that is helpful for understanding the need for, uses of, and issues with synthetic data. We will use a production library for synthetic modeling, so we will not address these concerns directly except as parameters to the model(s) we decide to use in the pipeline, but it is good to understand the kinds of trade-offs that are inherent in generative modeling for synthetic tabular data.

    \cite{DBLP:journals/corr/abs-1811-11264} introduces table GAN (TGAN), a GAN-based approach to generating synthetic tabular data. This model uses recurrent neural networks. Each of the generative modeling papers I reviewed have extensive coverage of how to handle concerns in the data. One interesting one to highlight, among a few, from this paper and demonstrating the type of complexity that goes into this type of modeling is the focus on multimodal continuous data. This is continuous data that has more than one value that stands out as more represented in the data, showing up as more than one peak when you plot a histogram of the data. To get around this, the researchers use Gaussian kernel density estimation (KDE) to estimate the distribution of the data. Each KDE can model one mode of the data. We then use Gaussian mixture methods (GMMs) to sample from more than one distribution -- each distribution coming from one of those built with the KDEs -- to recreate the actual distribution. We then sample from this GMM when we are generating a value for that feature, and this gets baked into the modeling, training, and generation processes. This is just one example to demonstrate why we are veering away from building our own generative models in this project.

    \cite{DBLP:journals/corr/abs-1907-00503}, by the same authors as the previous TGAN paper, provides more consideration for nuances and potential issues in the data along with adding a deep conditional GAN (DCGAN) as part of their new conditional tabular GAN (CTGAN) model. DCGAN generally is a GAN architecture that uses convolutional neural networks (CNNs) in both the generator and discriminator. One example of extra data considerations is that CTGAN handles severe imbalance in categorical features. We will not worry about the detail of this model either, but it is potentially one that we will use with the modeling library we introduce next

    \cite{sdv} is our project of choice for generative modeling. It is produced by the authors of the previous two papers above. Similar to those two papers, we will not dig too far into the workings of the libraries in this project. We mostly need a solution that can plug into the synthetic tabular data generative modeling step(s) in the pipeline. Note that both TGAN and CTGAN are dedicated repositories in the Synthetic Data Vault project. There is another main library -- sdv -- that provides easier-to-use versions of the models. There is also SDGym, a synthetic data evaluation library that we can lean on to determine how well our generative modeling is doing.

    With a sweep through some research papers out of the way, I did not find any papers directly applicable to what we are looking at in this project. So, the contribution that we provide is to start a discussion around data mining for IPEDS data along with generative modeling for IPEDS data. We build on the ideas in the papers above -- the general approaches to analysis in the earlier papers, and the use of synthetic tabular data for the latter papers.

\section{Proposed Work}
Dataset, tools
Main tasks
    - E.g., statistical analysis, visualization
    - E.g., data preprocessing, warehousing
    - E.g., specific questions, patterns to explore or model

\section{Evaluation}
Evaluation metrics
    - Effectiveness
    - Efficiency
Experimental setup
Methods to compare

\section{Discussion}
Project timeline
    - When to finish
    - Current status
    - Determine if you need to do less or can do more in the given time
Potential challenges
Alternative approaches and backup plan

\subsection{Questions for the Reader}
    \begin{itemize}
        \item Is adding in the synthetic data component too much? I've been exploring this is personal and school projects as well as using it at work, so it's fresh on my mind to keep testing out, but it may be overkill for this project.
    \end{itemize}

\section{Conclusion}
Conclusion
Project summary
Key findings (when finished)
Future work (when finished)

\bibliographystyle{ACM-Reference-Format}
\bibliography{msds_data_mining_project_report}

\appendix

\section{Technical Details}

    Need some details plz

\end{document}
\endinput