### 2025-02-08 report/msds_data_mining_project_report.tex
finish first pass of project report, before starting technical work

\documentclass[sigconf, authorversion, nonacm]{acmart}

\settopmatter{printacmref=false} % remove acm info that clutters things up for this project
\settopmatter{printfolios=true} % add page numbers

\citestyle{acmauthoryear}

\begin{document}

\title{Higher Education Synthesized}
\subtitle{Extending Higher Education Data with Synthetic Data at the Onset of the Demographic Cliff}

\renewcommand{\shortauthors}{Leakakos}

\begin{abstract}
    American higher education is coming to an impasse. Colleges and universities have seen enrollments grow as larger generations of Americans made their way through postsecondary education. Schools increased their academic offerings and student life activities. Faculty and staff members filled the needs for supporting increasing student enrollments. But there was an issue hiding behind all of this growth: what happens when student enrollments systematically fall?

    The United States has seen declines in high school graduates and, therefore, reduced college applicants. Depending on who you listen to, this ranges from a demographic cliff ushering in an emergency in higher education to a slower decline in higher education enrollments that will play out over years. It is unclear what effect this will have on the near future of higher education. One expected consequence of these declines is that we will see smaller amounts of higher education data relative to previous years.

    This project focuses on synthesizing Integrated Postsecondary Education Data System data from the National Center for Education Statistics. This data covers a vast range of information about higher education, and, while we will not look at student-level data --- one of the key areas for synthetic data after the demographic cliff hits --- we will start by generating synthetic data based on subsets of institution-related data. The goal is to show how we might address the diminishing data after the demographic cliff in order to continue higher education data analysis that relies on larger datasets.
\end{abstract}

\maketitle

\section{Introduction}
    American higher education saw growing enrollments throughout the latter half of the 1900s and moving into the 200s. This resulted in increases in the number of higher education institutions (HEIs) in the country, including public and private colleges and universities. These HEIs span from major research institutions down to small liberal arts schools. This has also resulted in increases in the number of faculty and staff to support growing enrollments. The general trend in higher education was growth.

    The issue is that this growth in enrollments was not to last. For years now, we have seen predicted and then realized declines in high school graduates along with the respective drop in higher education enrollments. Higher education's future was based on growth, but then the key aspect of that growth --- more students than in previous years --- fell away. Estimates for what comes next vary. Some see a demographic cliff and a looming emergency in which declining enrollments result in catastrophic cuts at and closures of HEIs. It is true that we have seen increasing numbers of HEI closures in recent years ~\cite{hechingercollegeclosures}. But others see the trend as a more gradual decline, giving higher education time to adapt and shrink accordingly. \cite{insidehighereddemographiccliff}

    There are many interesting research questions related to the future of higher education, appropriate and responsible responses to enrollment declines, and other aspects of this impending reality. There are also questions related to the shifting of costs from public funding to individual students, the shift in emphasis to higher education as workforce preparation, the modern publishing culture in higher education and how that affects the student experience, and more.

    But, this project will instead focus on a different aspect of declining higher education enrollments: reduced data. As student enrollments drop and HEIs close or shrink, we are likely to see reduced amounts of higher education data available for analysis, prediction, evaluation, and more. Not all analyses in higher education need larger amounts of data, but we want to explore how we could supplement real data with synthetic data in cases where we need more overall data. The risk is that, without sufficient data, we may not be able to support students and HEIs as well as we could with larger amounts of data.

    We will take a look at Integrated Postsecondary Education Data System (IPEDS) data from the National Center for Education Statistics (NCES), housed in the U.S. Department of Education's Institute of Education Sciences (IES). NCES gathers, prepares, and makes available data along with providing analysis on education data. IPEDS is a subset of NCES data that focuses on postsecondary or higher education data.

    We will focus on a subset of IPEDS data related to institutions, enrollments, and graduation rates. For each of these three domains, IPEDS offers multiple files covering different information, and it provides these for multiple years. Our goal will be to identify one or more analysis or modeling datasets that we can pull out of a larger extract-transform-load (ETL) pipeline with multiple repeated steps. We will focus on pulling IPEDS data, loading it into a relational database in its raw form, preprocessing it in the database with a series of automated processes, storing the analysis or modeling datasets in the database, and then using those datasets to synthesize new data.

   To address the issue of the amount of data available to us, we will leverage generative modeling to produce synthetic tabular data. The goal is to build a system that can demonstrate how we might leverage synthetic higher education data to fill in datasets in the future as they potentially shrink due to the issues described above.

    We are focusing on institution-level data from IPEDS, but future iterations of this work could pivot to generating synthetic student-level data.

    Synthetic data is ``artificially generated [data] that resemble[s] the actual data -- more precisely, having similar statistical properties'' ~\cite[p. 1]{decristofaro2024syntheticdatamethodsuse}. If we generate synthetic IPEDS data that shares the same properties as the real data, then we can combine the real and synthetic data, or, alternatively, we can create an indefinite amount of synthetic data, to increase the amount of data we have to work with.

    Building and training models to generate synthetic tabular data -- tabular meaning data in table form with rows and columns -- is notoriously difficult. We will leverage existing models and research for the synthetic generation aspect of the pipeline in particular. We will lean on the research in the Modeling Tabular Data Using Conditional GAN paper ~\cite{DBLP:journals/corr/abs-1907-00503} and the associated Python libraries they make available at the Synthetic Data Vault Project ~\cite{sdv}.

    While we will not dive too deep into the working of the generative models, it can be useful to understand the complexity in these and why we are leveraging existing implementations. The research paper focuses on their conditional tabular generative adversarial network (CTGAN) approach to generative modeling. This paper provides some examples of the difficulty of modeling for tabular data. Section 3 speaks to some of these issues:

    \begin{itemize}
        \item Mixed data types: tabular data can have features with continuous, nominal, ordinal, binary, or other data. Generative modeling for these datasets require handling mixtures of data types in one model.
        \item Non-Gaussian distributions: continuous data can be non-Gaussian, so attempts to standardize or min-max continuous features can result in issues such as vanishing or exploding gradients.
        \item Multimodal distributions: continuous features may have more than one mode, resulting in the need to model more than one mode with approaches such as kernel density estimation and Gaussian mixture models.
        \item Imbalance in categories for categorical features: categories that have smaller representation may not be modeled as accurately as those with larger representation, resulting in the need for extra steps to handle this disparity. \cite{DBLP:journals/corr/abs-1907-00503}
    \end{itemize}

    We do not need to go into the inner workings of these issues. Instead, they are meant to provide a snippet of the difficulties we would need to handle if building our own generative models. In order to remain focused on the pipeline in this project, we will, as said above, leverage existing generative modeling libraries available through the Synthetic Data Vault Project ~\cite{sdv}.

    Finally, the ultimate goal is to glean insights from the IPEDS data, real or synthesized. The synthetic data will ideally give us the same overall insights as the real data. For this task, we can turn to traditional data mining techniques: classification, clustering, association mining, rule mining, or others. If we can determine trends now, we may be able to utilize synthetic data in the future with dwindling real data to confirm continued trends or to identify changes in those trends.

    To summarize, this project will work with IPEDS data with its focus on higher education and will explore how we can synthesize more IPEDS data to address potential future lack-of-data issues due to declining higher education enrollments, and we will use these data to tackle standard data mining questions related but focused on higher education data.

\section{Related Work}
    My literature review did not return as much related to data mining in higher education as I expected to see, and I was unable to find any research closely related to what we explore in this project. The bulk of what I found is related to different evaluation methods for student learning. Examples are looking at adaptive learning --- learning that dynamically adjusts to students based on different input metrics --- and performance prediction --- trying to determine how a student will end up performing by, for example, graduation. While these are interesting, they do not focus on the type of data that we see in IPEDS.

    Note that some of these papers were already referenced in the introduction above.

    To start with, ~\cite{sdp} by the Harvard Strategic Data Project (SDP) fellowship program speaking about the importance of analytics in higher education data. There are larger questions that this paper raises about why higher education appears behind other areas in terms of leveraging data and more advanced analytics.

    Another example of an interesting paper that I want to highlight is ~\cite{Hassna02102023} that explores utilizing big data and analytics for higher education data. The catch is that this paper is from 2022, and it is talking about introducing big data and analytics in higher education, mirroring some of what we see from the SDP paper above in hinting that higher education may be behind other areas in terms of utilizing data. But, during the literature review, I checked out big data and data mining and related terms in other domains, and I did not tend to see papers talking about how big data and analytics have yet to be realized in those domains.

    Next is a handful of papers characterizing what I mostly found: research into specific uses of analytics and mining to hone in on an aspect of student evaluation. Since our focus on this project is on institution-level information with IPEDS data, we are zooming out quite a bit from what these papers cover, but these are useful in terms of getting bearings for research on higher education data mining, though none of these papers are overly useful for the specific use case in this project.

    First is ~\cite{studentbehavioranalytics} that talks about using higher education big data to evaluate student behavior. This paper repeats a similar claim to ~\cite{Hassna02102023} above in that higher education is missing work in big data.

    \cite{Okewu10112021} explores using artificial neural networks for higher education data mining. It provides a review of studies from 2010 through 2018 and related to education data mining focused on adaptive learning. Adaptive learning is the use of learner data to customize learning dynamically to those learners. So, we are looking at yet another narrower use case for a specific type of higher education learner data. This paper again mirrors some of the sentiments from earlier papers in that it hints that higher education data usage, and big data usage in particular, should be further along than where it is currently at (as of the paper's publishing in 2021), closer to where we see big data usage in other domains, but it is not.

    \cite{10.3389/fpsyg.2021.698490} investigates education data mining techniques for student performance prediction. There is not a ton more to add here aside from it being another narrower use case looking at student-level data, not the larger institution-level data that we explore in this project, but it is yet another example to learn from in terms of research using higher education data.

    \cite{ALDOWAH201913} digs into education data mining, similar to the last paper, but this time focused on learning analytics. It reviews studies in different domains related to learning analytics. The key takeaway is that it is yet another study that looks into different outcomes within specific areas of higher education data, but it is not close enough to what we are exploring in this project to lean on.

    Finally, ~\cite{https://doi.org/10.1155/2022/8924028} looks at predicting higher education student performance, typically at graduation, and it evaluates different predictive methods. Again, this is interesting for reviewing some current literature, but it will not apply to the work we do in this project.

    We will also include a quick review of a few papers related to synthetic data and synthetic tabular data generative modeling. We will utilize synthetic data, but we are not building it from scratch, so we will not dwell on the deep learning implementation details of these papers.

    \cite{decristofaro2024syntheticdatamethodsuse} provides an overview of synthetic data that is helpful for understanding the need for, uses of, and issues with synthetic data. We will use a production library for synthetic modeling, so we will not address these concerns directly except as parameters to the model(s) we decide to use in the pipeline, but it is good to understand the kinds of trade-offs that are inherent in generative modeling for synthetic tabular data.

    \cite{DBLP:journals/corr/abs-1811-11264} introduces table GAN (TGAN), a GAN-based approach to generating synthetic tabular data. This model uses recurrent neural networks. Each of the generative modeling papers I reviewed have extensive coverage of how to handle concerns in the data. One interesting one to highlight, among a few, from this paper and demonstrating the type of complexity that goes into this type of modeling is the focus on multimodal continuous data. This is continuous data that has more than one value that stands out as more represented in the data, showing up as more than one peak when you plot a histogram of the data. To get around this, the researchers use Gaussian kernel density estimation (KDE) to estimate the distribution of the data. Each KDE can model one mode of the data. We then use Gaussian mixture methods (GMMs) to sample from more than one distribution -- each distribution coming from one of those built with the KDEs -- to recreate the actual distribution. We then sample from this GMM when we are generating a value for that feature, and this gets baked into the modeling, training, and generation processes. This is just one example to demonstrate why we are veering away from building our own generative models in this project.

    \cite{DBLP:journals/corr/abs-1907-00503}, by the same authors as the previous TGAN paper, provides more consideration for nuances and potential issues in the data along with adding a deep conditional GAN (DCGAN) as part of their new conditional tabular GAN (CTGAN) model. DCGAN generally is a GAN architecture that uses convolutional neural networks (CNNs) in both the generator and discriminator. One example of extra data considerations is that CTGAN handles severe imbalance in categorical features. We will not worry about the detail of this model either, but it is potentially one that we will use with the modeling library we introduce next.

    \cite{sdv} is our project of choice for generative modeling. It is produced by the authors of the previous two papers above. Similar to those two papers, we will not dig too far into the workings of the libraries in this project. We mostly need a solution that can plug into the synthetic tabular data generative modeling step(s) in the pipeline. Note that both TGAN and CTGAN are dedicated repositories in the Synthetic Data Vault project. There is another main library -- sdv -- that provides easier-to-use versions of the models. There is also SDGym, a synthetic data evaluation library that we can lean on to determine how well our generative modeling is doing.

    With a sweep through some research papers out of the way, I did not find any papers directly applicable to what we are looking at in this project. So, the contribution that we provide is to start a discussion around data mining for IPEDS data along with generative modeling for IPEDS data. We build on the ideas in the papers above -- the general approaches to analysis in the earlier papers, and the use of synthetic tabular data for the latter papers.

\section{Proposed Work}
    For me, at this stage, this is the section that I am looking to narrow down a bit (and I am open to suggestions from readers). I have a couple of ideas planned for the work in this project, and I am looking to lean into my background with the architecting of data systems to build out a wider pipeline.

    We are exploring IPEDS data. We went over the value of that data earlier. To snag the data, we will use some Python scripts to download the data from the IPEDS website and store it in a relational database. This database will run locally in Docker, and the configuration files needed to set up the database will exist in this repository, both for the Docker setup as well as for the database work. IPEDS data already comes in a relatively relational form, so there may not be too much work for the initial load into the database.

    From there, we may leverage the database further to provide preprocessing in the database. One easy way to do this is to set up insert triggers to handle the first pass of standardizing the data into a consistent format or doing some early combining and mutating of data before looking at larger exploratory data analysis (EDA). This may work, but we may also want to pull the data out of the database to go through a more standard EDA. But, there is an option here to mimic what we may need in a larger production pipeline with leveraging database tech at this stage.

    Next we go through a full EDA process, this time with a mind towards how we may want to use the data later in the pipeline rather than the earlier EDA based around simple cleanup. This step will happen outside of the database, so we will look at Python scripts here. At first, this is likely easier with the read-eval-print loop (REPL) functionality of a notebook, though we do not want notebooks in our final pipeline until much closer to the end in production. We need the automation power in scripts instead of notebooks.

    After this round of EDA and any further preprocessing we glean from it, we can then think about warehousing. We will shift the data back into the warehouse, likely just matching the new form of the data to what we need for modeling, but we can decide later on the form of the data in the warehouse. For the scope of this project, we will likely end up with a bespoke warehouse built to immediate modeling or analysis needs rather than a broader warehouse set up in something like a full star schema.

    The next step is to explore the outputs of the pipeline. Two directions stand out to me. One is modeling. Especially when working with larger datasets such as those we deal with in data mining, prediction and clustering jump out as obvious candidates, with clustering typically being much more complicated to set up. The second direction is to explore some data mining-specific approaches and algorithms such as similarity measures, apriori, FP-growth, association or rule mining, or others. Honestly, at this stage, any of these could be valid, and I prefer to start working with the data some before deciding on outputs at this stage, so this is left purposely broad.

    One piece I am pulling to the end here is the synthetic data component. I am already wondering if adding in synthetic data is too much. I have been exploring synthetic data in personal and school projects as well as using it at work, so it is fresh on my mind, and I would like to keep testing it out, but it may be overkill for this project. If we do keep synthetic data, then it makes sense to simplify some of the rest of the pipeline to keep more of the scope focused on the synthetic data modeling. This means we probably focus more at the end of the pipeline on evaluating the quality of the synthetic data and demoing how we would pull insights out of the synthetic dataset, but we probably look at just one approach for that last piece instead of going through a heavy refinement of the other outputs.

    The areas I am keeping an eye on here are the extra loop of preprocessing in the database, the inclusion of synthetic data, and narrowing down the final outputs. The ideas in this section are fine for now, but we need to make these more concrete as we begin digging into the data.

    I have spent most of the time here focusing on the technical side of the work. I need to dig into the IPEDS data before feeling more comfortable with the questions in terms of meaning that we will explore, but I expect those to center around identifying patterns in clusters of institutions based on institution attributes, enrollment trends, and graduation trends, possibly even dropping either enrollment or graduation data in favor of simplifying the data scope for right now.

\section{Evaluation}
    Until we decide to drop synthetic data, if we do make that decision, the first evaluations we will explore are how similar the synthetic data is to the real data. We will explore the evaluation metrics in the Synthetic Data Vault Project. A couple of candidates that stand out are per-attribute comparisons and machine learning score. Per-attribute comparisons check the same feature in the real and synthetic datasets, calculate metrics such as mean or standard deviation, and determine the quality of the synthetic data based on how close these measures are across the datasets.

    Machine learning score trains classifiers on the real and synthetic datasets and compares the performance of those models. For example, if we use both classifiers to check something like accuracy or F1 score on a held-out test set, we can evaluate how similar the synthetic data is to the real data based on comparing those scores.

    We can also look at more advanced evaluation metrics. One that has stood out to me predicting each feature in a dataset with all of the other features. If we find the predictive power in the real and synthetic datasets are similar, then, like above, that can give us a sense of how similar the datasets themselves are.

    There are any number of other evaluative metrics for synthetic data. We will explore what is available with the project we use for synthetic data and will clarify here as we learn more during experimentation.

    Moving on from synthetic data, until we explore the IPEDS data further, I am not sure exactly what we will end up modeling or testing as output. But there are some general contenders that we could likely look at.

    For supervised modeling outputs, we can look at standard metrics such as accuracy, precision, F1 score, and more for classification problems. We can use a metric like mean squared error (MSE) to measure predictive model output. These evaluations give us firm numbers we can use to determine how well our modeling is doing.

    But, I am actually more intrigued by seeing what we can find with unsupervised learning. Admittedly, the outputs for unsupervised learning can be more difficult to work with, especially at this early stage. Clustering is the first unsupervised modeling that comes to mind. I often think of clustering as a much more advanced version of EDA. We are not really sure what patterns exist in the data, so we supercharge standard EDA and see what patterns we uncover when leveraging machine learning. But we do not have a clean evaluation metric with clustering. We will need to interactively explore the clustering outputs, at least at first. At some point, we can determine a set number of clusters and pivot the clustering into a classification approach, but the more interesting piece here for me is the clustering itself.

    I am on the fence about leaning more into the synthetic data approach in which we would spend a lot more time evaluating the quality of the synthetic data. With synthetic data, I still would like to test out one of the analytical pieces at the end of the pipeline, but maybe we stick to something simpler like a clearer classification problem. If we instead drop synthetic data, then we have more room to tackle something like unsupervised learning and the complexity inherent in it.

    While I feel like there are loads of evaluative methods to explore, I am also interested in looking at some of the algorithms from the second data mining course. I have leveraged things like similarity scores in other projects, but I have not experimented with, for instance, FP growth, and it would be interesting to possibly pick a brand new algorithm to plop at the end of the pipeline and figure out what it means to evaluate that algorithm and what it reveals to us.

    So, at this stage, this section is more of a brainstorming of some of the evaluative approaches. This will get refined down in the next iteration after exploring the IPEDS data a bit and determining if we will keep or drop the synthetic data component of the pipeline.

\section{Discussion}
    I tend to drag out projects, taking the opportunity to add increased complexity and explore new items I am not as familiar with. For this project, part of me wants to focus on sticking a bit more in-bounds in terms of my comfort level with data pipelines, only adding in a little extra complexity and seeing if it is feasible to build the entire pipeline out in a shorter amount of time. While spending forever on projects is fun and educational, I do risk getting too comfortable with taking my time. I could benefit from descoping and focusing on getting smaller chunks working earlier and then deciding if it is worth expanding out from there.

    In that vein, I am hoping that the refinement in going from the first to second draft of this report, spending some time to explore the technical and data sides of the project in between, will help reveal scope from this draft that can be cut. It would be ideal if I could use a three or four week timeline and adjust the project to fit within that, but it does mean that what I have laid out above is probably too ambitious for that timelines. But, after writing this all out, the scope does feel pretty big at the moment.

    Currently I have identified some IPEDS data, but I have not written any code or set up the database. My approach will be to get the data through to at least the initial EDA and preprocessing before starting to evaluate the scope after that. I feel comfortable with the earlier and more technical steps in terms of time, so my question marks sit with the longer EDA, the modeling, and, if included, the synthetic data portion.

    The main challenge I see is that the data ends up being much more complicated to work with than expected, so I will need to shift the scope down even further to prioritize working with difficult data. That would not be a bad way for the project to pivot, but it is not what I am planning on at this stage. If the data turns out to be friendly enough to work with, then I think the next big challenge will be figuring out the right analysis or modeling in terms of the meaning of the data. The ideas above are all interesting from a technical standpoint, but the meaning of what we dig into needs to drive why we are using those approaches.

    The general backup plan is to limit each stage of the pipeline to maybe one or two pieces. This may mean we do some simple EDA along with a correlation matrix, we look at a clearer regression or classification problem, or we leverage synthetic data but trim the input dataset to only a few features. All of these are tactics for trimming the scope, likely the biggest risk for the project at the moment.

\section{Conclusion}
    Not much to conclude at the moment. We will fill this in as we start doing the data work of the project. What I expect to see here is that we trimmed down the IPEDS data, scoped the use case for the project to be a bit narrower, decided to explore synthetic data, and picked a simpler overall output. For a surprise finish, we may end up deciding to pull some data out and run it through a separate business intelligence tool like Tableau with the idea that end users of the production pipeline may want to leverage tools such as that for their purposes.

    I expect the next steps to end up being that we incorporate more IPEDS data to answer a broader range of questions. I do not expect radical changes in the pipeline architecture in next iterations of the project, but I do expect additions and expansions.

\bibliographystyle{ACM-Reference-Format}
\bibliography{msds_data_mining_project_report}

\appendix

\section{Technical Details}
    Need some details plz. Leaving this as a placeholder to pull the technical details out of the narrative above but still keep it available for readers.

\end{document}
\endinput